Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-07-03 draftfinalreport_v4.tex
text revisions based on changes in naive %
Changed the wording to reflect the altered % and numbers to vaccinate.

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand, 2013} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 17\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve to measles, it was estimated that approximately 54\% of those would require additional vaccination to ensure measles does not persist and outbreaks do not occur, leading to an additional 402,010 effective vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially immunised population) values often include or exceed one, including 2013--2014 outbreaks that were ongoing at the time of analysis. As analysed from data until 12 June 2014, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of the first 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost of those measles cases that attend hospital is estimated to be \$1,877 per attending case. Approximately 17\% of measles cases attend hospital.
\item The mean public health service cost per case (not including hospitalisation costs) is \$1,765.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccination-related costs required to exceed approximately \$1,117 per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs together.
\item The B/C ratio differed across DHBs with those with large populations and more na\"{\i}ve in absolute numbers most likely to benefit from immunisation campaigns. 
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 54\% of the currently 17\% na\"{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@

\vspace{5mm} %5mm vertical space
\begin{table}
\begin{center}
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 12:59:00 2015
\begin{tabular}{lr}
  \hline
 & Pr($>$F) \\ 
  \hline
Age & 0.0000 \\ 
  Ethnicity & 0.0141 \\ 
  NZDep & 0.0102 \\ 
  Age:Ethnicity & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure. Pr($>$F) is p-value, the statistical significance of the result}
\label{table:anova}
\end{center}
\end{table}

\vspace{5mm} %5mm vertical space
\begin{table}
\begin{center}
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 12:59:00 2015
\begin{tabular}{rrrr}
  \hline
 & Estimate & Std. Error & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.16 & 0.13 & 0.00 \\ 
  Age2-4 & -1.04 & 0.20 & 0.00 \\ 
  Age5-17 & -1.00 & 0.14 & 0.00 \\ 
  Age18-24 & -1.72 & 0.19 & 0.00 \\ 
  Age25+ & -3.17 & 0.16 & 0.00 \\ 
  EthnicityAsian & 0.25 & 0.32 & 0.44 \\ 
  EthnicityMaori & 0.21 & 0.21 & 0.33 \\ 
  EthnicityPacific & 1.22 & 0.21 & 0.00 \\ 
  NZDep4-7 & -0.26 & 0.09 & 0.01 \\ 
  NZDep8-10 & -0.30 & 0.10 & 0.01 \\ 
  Age2-4:EthnicityAsian & -2.31 & 1.33 & 0.09 \\ 
  Age5-17:EthnicityAsian & -1.25 & 0.46 & 0.01 \\ 
  Age18-24:EthnicityAsian & -1.02 & 0.57 & 0.08 \\ 
  Age25+:EthnicityAsian & 0.23 & 0.43 & 0.59 \\ 
  Age2-4:EthnicityMaori & -0.10 & 0.36 & 0.78 \\ 
  Age5-17:EthnicityMaori & -0.10 & 0.25 & 0.68 \\ 
  Age18-24:EthnicityMaori & -0.57 & 0.44 & 0.20 \\ 
  Age25+:EthnicityMaori & -1.03 & 0.51 & 0.05 \\ 
  Age2-4:EthnicityPacific & -0.98 & 0.47 & 0.04 \\ 
  Age5-17:EthnicityPacific & -1.57 & 0.33 & 0.00 \\ 
  Age18-24:EthnicityPacific & -1.04 & 0.50 & 0.05 \\ 
  Age25+:EthnicityPacific & -0.66 & 0.43 & 0.13 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-12 June 2014. Estimate: negative values indicate rates are lower for this category compared to the base category, which does not appear (e.g. Age 5--17 has a lower measles rate per capita compared to the 0--1 age class that is not shown), and positive values indicate the rates are greater than the baseline rate; Std. error is the standard error, a measure of uncertainty in the estimates; and Pr(>|t|) is the p-value, the statistical significance of the result.}
\label{table:regression_riskfactors}
\end{center}
\end{table}

@@

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{naive_allPop_yob.pdf}
\end{center}
\caption{New Zealand population by year of birth and estimated numbers of na\"{\i}ve people in each age class using national immunity data (\autoref{table:sero}) (top) and the number of cases by year of birth (bottom).}
\label{fig:naive_yob}
\end{figure}

@@

The majority of the na\"{\i}ve among the general New Zealand population are in their first years of life (\autoref{fig:naive_yob}). However, the distribution of na\"{\i}ve at a national level shows that the recent MMR vaccination schemes are reducing the proportions of na\"{\i}ve population in the 3--5 year old age group to greater levels than in older young people  (\autoref{fig:naive_yob}). The distribution of the numbers of na\"{\i}ve and the total na\"{\i}ve populations per DHB, assuming national immunisation and immunity rates are representative, are given in the following figures in the appendix (\autoref{sec:app}): 
\begin{itemize}[noitemsep,nolistsep]
\item Northland: \autoref{fig:Northland}
\item Waitemata: \autoref{fig:Waitemata}
\item Auckland: \autoref{fig:Auckland}
\item Counties Manukau: \autoref{fig:Counties_Manukau}
\item Waikato: \autoref{fig:Waikato}
\item Lakes: \autoref{fig:Lakes}
\item Bay of Plenty: \autoref{fig:BayofPlenty}
\item Tairawhiti: \autoref{fig:Tairawhiti}
\item Taranaki: \autoref{fig:Taranaki}
\item Hawke's Bay: \autoref{fig:HawkesBay}
\item Whanganui: \autoref{fig:Whanganui}
\item Mid-Central: \autoref{fig:Midcentral}
\item Hutt: \autoref{fig:Hutt}
\item Capital and Coast: \autoref{fig:CapitalandCoast}
\item Wairarapa: \autoref{fig:Wairarapa}
\item Nelson Marlborough: \autoref{fig:NelsonMarlborough}
\item West Coast: \autoref{fig:WestCoast}
\item Canterbury: \autoref{fig:Canterbury}
\item South Canterbury: \autoref{fig:SouthCanterbury}
\item Southern: \autoref{fig:Southern}
\end{itemize}

@@

The distribution of the measles cases (\autoref{fig:ageandvac}) appears to reflect the distribution of na\"{\i}ve individuals in the population (\autoref{fig:naive_yob}). However, there is some discrepancy between the year of birth of cases (\autoref{fig:yearandvac}) and the proportions na\"{\i}ve (\autoref{fig:naive_yob}), which may reflect other processes like spatial effects. The vaccination history of the cases suggests that lack of vaccination cover is the main contributor to the outbreak, though it is noticeable that a number ($>16\%$) of cases had been vaccinated at least once. However, the majority of these received only one vaccination (82\%), and were vaccinated when they were young (44\% of those cases vaccinated once were 15 months or younger, see \autoref{fig:vaccstat} and \autoref{fig:vaccstat_close}).

@@

To use the results from our modelling exercise to help inform the appropriate measles vaccination coverage, we use \autoref{eq:prop}. The proportion of the na\"{i}ve population requiring additional vaccination to make $R_v$ < 1 was very similar across all DHBs (54--55\% na\"{i}ve), with a national average of approximately 54\% (\autoref{table:attack}). These additional vaccination numbers can be calculated in a number of different ways and the benefits of vaccination vary, as discussed in the benefit--cost section (\autoref{sub:cost_benefit}). However, they require differing numbers of vaccinations per DHB. Estimates for West Coast, Wairarapa, and Tairawhiti, for example, are fewer than 5000 (\autoref{table:attack}). The estimated numbers to vaccinate in the Auckland area, however, are higher for each DHB, with those estimated for Waitemata, Auckland, and Counties Manukau DHBs all over 40,000  (\autoref{table:attack}). Note that there is also evidence for vaccine failure (\autoref{fig:vaccstat} and \autoref{fig:vaccstat_close}) and pockets of susceptible populations \citep{glass4} may require greater vaccination effort than these estimates and this is considered further in the \autoref{sec:cb}.

@@

\begin{table}
\small
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 12:59:00 2015
\begin{tabular}{lrrrrr}
  \hline
DHB & Population & Naive & Outbreak & Vaccination & PC \\ 
  \hline
Auckland & 436350 & 74410 & 10662 & 40320 & 0.54 \\ 
  Bay of Plenty & 206000 & 35150 & 5041 & 19056 & 0.54 \\ 
  Canterbury & 482180 & 82195 & 11770 & 44525 & 0.54 \\ 
  Capital and Coast & 283700 & 48368 & 6928 & 26204 & 0.54 \\ 
  Counties Manukau & 469300 & 80484 & 11633 & 43820 & 0.54 \\ 
  Hawke's Bay & 151700 & 25934 & 3730 & 14082 & 0.54 \\ 
  Hutt Valley & 138380 & 23864 & 3478 & 13053 & 0.55 \\ 
  Lakes & 98196 & 16854 & 2439 & 9182 & 0.54 \\ 
  MidCentral & 162560 & 27397 & 3855 & 14697 & 0.54 \\ 
  Nelson Marlborough & 137000 & 23502 & 3398 & 12799 & 0.54 \\ 
  Northland & 151690 & 25902 & 3719 & 14051 & 0.54 \\ 
  South Canterbury & 55620 & 9444 & 1344 & 5099 & 0.54 \\ 
  Southern & 297420 & 50047 & 7025 & 26811 & 0.54 \\ 
  Tairawhiti & 43650 & 7473 & 1077 & 4063 & 0.54 \\ 
  Taranaki & 109750 & 18842 & 2728 & 10268 & 0.54 \\ 
  Waikato & 359310 & 61138 & 8730 & 33067 & 0.54 \\ 
  Wairarapa & 41112 & 7005 & 1002 & 3793 & 0.54 \\ 
  Waitemata & 525550 & 90522 & 13170 & 49463 & 0.55 \\ 
  West Coast & 32151 & 5554 & 812 & 3042 & 0.55 \\ 
  Whanganui & 60120 & 10218 & 1457 & 5521 & 0.54 \\ 
  TOTAL & 4353189 & 742103 & 106278 & 402010 & 0.54 \\ 
   \hline
\end{tabular}\caption{Epidemic size calculations. \textit{Population}: DHB Population, Statistics NZ 2013;  	
\textit{Na\"{\i}ve}: DHB na\"{\i}ve population $\left(x_0\times\text{Size}\right)$ (see \autoref{table:sero} for immunity estimates);
\textit{Outbreak}: Number infected in DHB in an outbreak of measles in the absence of any further vaccination assuming homogeneous mixing across DHB $\left(\Pe\right)$. Note outbreaks of these size lead to immunity rates that leave $\Rr_V$ below one until enough susceptibles are accumulated;	
\textit{Vaccination}:	Number of the currently na\"{\i}ve to be immunised in DHB to reduce $\Rr_V$ below one  $\left(\left(x_0-1/\Ro\right)\times\text{Size}\right)$; \textit{PC}: The proportion of the currently na\"{\i}ve population to vaccinate. These simulations assume homogeneous mixing. See \autoref{sub:mod_meth} for equations and further details on calculations.
}
\label{table:attack}
\end{table}

\subsection{Summary of modelling}
\begin{itemize}
\item Regular introductions of measles pose an ongoing threat to New Zealand's efforts to eliminate measles (also see \autoref{sub:risk_analyses}).
\item The reproduction number for measles in the partially immune New Zealand population is often close to and exceeding one, suggesting increased population level immunity is required to ensure prevention of measles persistence following importation.
 \item The proportion of the currently na\"{\i}ve population requiring additional and effective vaccination in New Zealand required to push $R_v$ below one ranges from 17\% to 34\% among DHBs, and is 54\% at the national level (approximately 402,010 vaccinations).
\end{itemize}

>>>>

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand, 2013} report was of high quality and contained substantial information and useful analyses.
\item The regression analyses suggest that age is a particularly strong risk factor for measles.  Increased wealth (NZDep 1--3) was a risk factor and people of European ethnicity make up the majority of cases. However, our analyses also highlight other groups that are at greater risk of measles on a per capita basis. Pacific people are at greater risk per capita. However, this was driven by young Pacific children, as 2--24 year old Pacific people are less at risk than European and Maori, as are 5--17 year old Asian children.
\item Of the 17\% of the overall population of New Zealand that are currently considered to be immunologically na\"{\i}ve to measles, it was estimated that approximately 54\% of those would require additional vaccination on top of current childhood vaccination rates to ensure measles does not persist and outbreaks do not occur, leading to an additional 402,010 effective vaccinations nationally.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe. Peak overall overseas travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travel is otherwise out of phase, with peak travel for New Zealanders during winter, and non--New Zealanders during summer.
\item Analyses of outbreak data suggest that measles basic reproduction number ($R_v$, the number of secondary infections in a partially immunised population) values often include or exceed one, including 2013--2014 outbreaks that were ongoing at the time of analysis. As analysed from data until 12 June 2014, the last outbreak to end in 2014 had an $R_v$ well above 1 (mean estimate 3.66, range 3.4--3.92). The range of estimates for $R_v$ between 2009--2014 was 0.18--3.92. These analyses suggest improved vaccination is a requisite to prevent measles persisting in the population and that large outbreaks could occur.
\item The cost of the first 187 measles cases in 2014 is estimated to be approximately \$1,041,186 due to earnings lost relating to cases and contacts, case management and hospitalisation costs.
\item The mean wage loss per measles case is estimated to be approximately \$839.
\item The mean cost of those measles cases that attend hospital is estimated to be \$1,877 per attending case. Approximately 17\% of measles cases attend hospital.
\item The mean public health service cost per case (not including hospitalisation costs) is \$1,765.
\item The average number of contacts requiring quarantine was 2.11, requiring 7.3 days of quarantine on average, at a cost of \$170 per day.
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>$1), with vaccination-related costs required to exceed approximately \$1,117 per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs together.
\item The B/C ratio differed across DHBs with those with large populations and more na\"{\i}ve in absolute numbers most likely to benefit from immunisation campaigns. 
\item Following catch up vaccination to ensure $R_v$ is < 1 measles introductions to New Zealand with a median size of 2 cases were predicted by simulation. Thus, typically individual cases and secondary cases will be the expectation. However, the mean outbreak size of 61 cases was predicted, because outbreak sizes of tens, hundreds, and very rarely up to many thousands of cases are predicted among the 1000 simulations following importation, despite $R_v$ being one and the outbreak predicted to die out. Thus, increased vaccination beyond the 54\% of the currently 17\% na\"{\i}ve population required may be useful to prevent these rare but costly events.
\end{itemize}

@@

\vspace{5mm} %5mm vertical space
\begin{table}
\begin{center}
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 14:02:33 2015
\begin{tabular}{lr}
  \hline
 & Pr($>$F) \\ 
  \hline
Age & 0.0000 \\ 
  Ethnicity & 0.0141 \\ 
  NZDep & 0.0102 \\ 
  Age:Ethnicity & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure. Pr($>$F) is p-value, the statistical significance of the result}
\label{table:anova}
\end{center}
\end{table}

\vspace{5mm} %5mm vertical space
\begin{table}
\begin{center}
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 14:02:33 2015
\begin{tabular}{rrrr}
  \hline
 & Estimate & Std. Error & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.16 & 0.13 & 0.00 \\ 
  Age2-4 & -1.04 & 0.20 & 0.00 \\ 
  Age5-17 & -1.00 & 0.14 & 0.00 \\ 
  Age18-24 & -1.72 & 0.19 & 0.00 \\ 
  Age25+ & -3.17 & 0.16 & 0.00 \\ 
  EthnicityAsian & 0.25 & 0.32 & 0.44 \\ 
  EthnicityMaori & 0.21 & 0.21 & 0.33 \\ 
  EthnicityPacific & 1.22 & 0.21 & 0.00 \\ 
  NZDep4-7 & -0.26 & 0.09 & 0.01 \\ 
  NZDep8-10 & -0.30 & 0.10 & 0.01 \\ 
  Age2-4:EthnicityAsian & -2.31 & 1.33 & 0.09 \\ 
  Age5-17:EthnicityAsian & -1.25 & 0.46 & 0.01 \\ 
  Age18-24:EthnicityAsian & -1.02 & 0.57 & 0.08 \\ 
  Age25+:EthnicityAsian & 0.23 & 0.43 & 0.59 \\ 
  Age2-4:EthnicityMaori & -0.10 & 0.36 & 0.78 \\ 
  Age5-17:EthnicityMaori & -0.10 & 0.25 & 0.68 \\ 
  Age18-24:EthnicityMaori & -0.57 & 0.44 & 0.20 \\ 
  Age25+:EthnicityMaori & -1.03 & 0.51 & 0.05 \\ 
  Age2-4:EthnicityPacific & -0.98 & 0.47 & 0.04 \\ 
  Age5-17:EthnicityPacific & -1.57 & 0.33 & 0.00 \\ 
  Age18-24:EthnicityPacific & -1.04 & 0.50 & 0.05 \\ 
  Age25+:EthnicityPacific & -0.66 & 0.43 & 0.13 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-12 June 2014. Estimate: negative values indicate rates are lower for this category compared to the base category, which does not appear (e.g. Age 5--17 has a lower measles rate per capita compared to the 0--1 age class that is not shown), and positive values indicate the rates are greater than the baseline rate; Std. error is the standard error, a measure of uncertainty in the estimates; and Pr(>|t|) is the p-value, the statistical significance of the result.}
\label{table:regression_riskfactors}
\end{center}
\end{table}

@@

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{naive_allPop_yob.pdf}
\end{center}
\caption{New Zealand population by year of birth and estimated numbers of na\"{\i}ve people in each age class using national immunity data (\autoref{table:sero}) (top) and the number of measles cases by year of birth from 2007--2014 (see \autoref{fig:ageandvac}) (bottom).}
\label{fig:naive_yob}
\end{figure}

@@

The majority of the na\"{\i}ve among the general New Zealand population are in their first years of life (\autoref{fig:naive_yob}). However, the distribution of na\"{\i}ve at a national level shows that the recent MMR vaccination schemes are reducing the proportions of na\"{\i}ve population in the younger age groups ($\geq$ 2006 year of birth) than in older young people  (\autoref{fig:naive_yob}). The distribution of the numbers of na\"{\i}ve and the total na\"{\i}ve populations per DHB, assuming national immunisation and immunity rates are representative, are given in the following figures in the appendix (\autoref{sec:app}): 
\begin{itemize}[noitemsep,nolistsep]
\item Northland: \autoref{fig:Northland}
\item Waitemata: \autoref{fig:Waitemata}
\item Auckland: \autoref{fig:Auckland}
\item Counties Manukau: \autoref{fig:Counties_Manukau}
\item Waikato: \autoref{fig:Waikato}
\item Lakes: \autoref{fig:Lakes}
\item Bay of Plenty: \autoref{fig:BayofPlenty}
\item Tairawhiti: \autoref{fig:Tairawhiti}
\item Taranaki: \autoref{fig:Taranaki}
\item Hawke's Bay: \autoref{fig:HawkesBay}
\item Whanganui: \autoref{fig:Whanganui}
\item Mid-Central: \autoref{fig:Midcentral}
\item Hutt: \autoref{fig:Hutt}
\item Capital and Coast: \autoref{fig:CapitalandCoast}
\item Wairarapa: \autoref{fig:Wairarapa}
\item Nelson Marlborough: \autoref{fig:NelsonMarlborough}
\item West Coast: \autoref{fig:WestCoast}
\item Canterbury: \autoref{fig:Canterbury}
\item South Canterbury: \autoref{fig:SouthCanterbury}
\item Southern: \autoref{fig:Southern}
\end{itemize}

@@

The distribution of the measles cases (\autoref{fig:ageandvac}) appears to reflect the distribution of na\"{\i}ve individuals in the population (\autoref{fig:naive_yob}) to some degree. However, there is some discrepancy between the year of birth of cases (\autoref{fig:yearandvac}) and the proportions na\"{\i}ve (\autoref{fig:naive_yob}), which may reflect other processes like spatial effects in contact rates and therefore measles transmission. The vaccination history of the cases suggests that lack of vaccination cover is the main contributor to the outbreak, though it is noticeable that a number ($>16\%$) of cases had been vaccinated at least once. However, the majority of these received only one vaccination (82\%), and were vaccinated when they were young (44\% of those cases vaccinated once were 15 months or younger, see \autoref{fig:vaccstat} and \autoref{fig:vaccstat_close}).

@@

To use the results from our modelling exercise to help inform the appropriate measles vaccination coverage, we use \autoref{eq:prop}. The proportion of the na\"{i}ve population requiring additional vaccination over and above the current childhood vaccination coverage to make $R_v$ < 1 was very similar across all DHBs (54--55\% na\"{i}ve), with a national average of approximately 54\% (\autoref{table:attack}). These additional vaccination numbers can be calculated in a number of different ways and the benefits of vaccination vary, as discussed in the benefit--cost section (\autoref{sub:cost_benefit}). However, they require differing numbers of vaccinations per DHB. Estimates for West Coast, Wairarapa, and Tairawhiti, for example, are fewer than 5000 (\autoref{table:attack}). The estimated numbers to vaccinate in the Auckland area, however, are higher for each DHB, with those estimated for Waitemata, Auckland, and Counties Manukau DHBs all over 40,000  (\autoref{table:attack}). Note that there is also evidence for vaccine failure (\autoref{fig:vaccstat} and \autoref{fig:vaccstat_close}) and pockets of susceptible populations \citep{glass4} may require greater vaccination effort than these estimates and this is considered further in the \autoref{sec:cb}.

@@

\begin{table}
\small
% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Fri Jul 03 14:02:33 2015
\begin{tabular}{lrrrrr}
  \hline
DHB & Population & Naive & Outbreak & Vaccination & PC \\ 
  \hline
Auckland & 436350 & 74410 & 10662 & 40320 & 0.54 \\ 
  Bay of Plenty & 206000 & 35150 & 5041 & 19056 & 0.54 \\ 
  Canterbury & 482180 & 82195 & 11770 & 44525 & 0.54 \\ 
  Capital and Coast & 283700 & 48368 & 6928 & 26204 & 0.54 \\ 
  Counties Manukau & 469300 & 80484 & 11633 & 43820 & 0.54 \\ 
  Hawke's Bay & 151700 & 25934 & 3730 & 14082 & 0.54 \\ 
  Hutt Valley & 138380 & 23864 & 3478 & 13053 & 0.55 \\ 
  Lakes & 98196 & 16854 & 2439 & 9182 & 0.54 \\ 
  MidCentral & 162560 & 27397 & 3855 & 14697 & 0.54 \\ 
  Nelson Marlborough & 137000 & 23502 & 3398 & 12799 & 0.54 \\ 
  Northland & 151690 & 25902 & 3719 & 14051 & 0.54 \\ 
  South Canterbury & 55620 & 9444 & 1344 & 5099 & 0.54 \\ 
  Southern & 297420 & 50047 & 7025 & 26811 & 0.54 \\ 
  Tairawhiti & 43650 & 7473 & 1077 & 4063 & 0.54 \\ 
  Taranaki & 109750 & 18842 & 2728 & 10268 & 0.54 \\ 
  Waikato & 359310 & 61138 & 8730 & 33067 & 0.54 \\ 
  Wairarapa & 41112 & 7005 & 1002 & 3793 & 0.54 \\ 
  Waitemata & 525550 & 90522 & 13170 & 49463 & 0.55 \\ 
  West Coast & 32151 & 5554 & 812 & 3042 & 0.55 \\ 
  Whanganui & 60120 & 10218 & 1457 & 5521 & 0.54 \\ 
  TOTAL & 4353189 & 742103 & 106278 & 402010 & 0.54 \\ 
   \hline
\end{tabular}\caption{Epidemic size calculations. \textit{Population}: DHB Population, Statistics NZ 2013;  	
\textit{Na\"{\i}ve}: DHB na\"{\i}ve population $\left(x_0\times\text{Size}\right)$ (see \autoref{table:sero} for immunity estimates);
\textit{Outbreak}: Number infected in DHB in an outbreak of measles in the absence of any further vaccination assuming homogeneous mixing across DHB $\left(\Pe\right)$. Note outbreaks of these size lead to immunity rates that leave $\Rr_V$ below one until enough susceptibles are accumulated;	
\textit{Vaccination}:	Number of the currently na\"{\i}ve to be immunised in DHB to reduce $\Rr_V$ below one  $\left(\left(x_0-1/\Ro\right)\times\text{Size}\right)$; \textit{PC}: The proportion of the currently na\"{\i}ve population to vaccinate. These simulations assume homogeneous mixing. See \autoref{sub:mod_meth} for equations and further details on calculations.
}
\label{table:attack}
\end{table}

\subsection{Summary of modelling}
\begin{itemize}
\item Regular introductions of measles pose an ongoing threat to New Zealand's efforts to eliminate measles (also see \autoref{sub:risk_analyses}).
\item The reproduction number for measles in the partially immune New Zealand population is often close to and exceeding one, suggesting increased population level immunity is required to ensure prevention of measles persistence following importation.
 \item The proportion of the currently na\"{\i}ve population requiring additional and effective vaccination in New Zealand required to push $R_v$ below one is 54\% at the national level (approximately 402,010 vaccinations), but varies from a few thousand to >40,000 at the DHB level.
\end{itemize}
